Synthesizing Symmetric Lenses

### 2018-07-12 papers/SymmetricLenses/main.tex
merge main

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing data to be transformed in two directions.  Since their
introduction by~\citet{Focal2005-long2}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much richer class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long2}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.

>>>>

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing data to be transformed in two directions.  Since their
introduction by~\citet{Focal2005-long}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much richer class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.
